Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization

### 2019-05-09 paper/paper.Rmd
small revisions

---
title: 'Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization
  After Passive Visuoproprioceptive Discrepancy Training'
author:
- affiliation:
  - YORK
  - MANSOURA
  name: Ahmed A. Mostafa
- affiliation: YORK
  corresponding: thartbm@gmail.com
  email: thartbm@gmail.com
  name: Bernard Marius 't Hart
- affiliation: YORK
  email: deniseh@yorku.ca
  name: Denise Y.P. Henriques
address:
- address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
    Canada
  code: YORK
- address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
  code: MANSOURA
output:
  html_document:
    df_print: paged
  pdf_document: default
  rticles::plos_article: default
  word_document: default
csl: plos.csl
fig_caption: yes
header-includes:
- \usepackage{threeparttable}
- \usepackage[table]{xcolor}
- \usepackage{graphicx}
- \usepackage{tabularx}
- \usepackage{longtable}
- \usepackage{siunitx}
bibliography: bibliography.bib
abstract: |
  An accurate estimate of limb position is necessary for movement planning, before and after motor learning. Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but in studies and theories on motor adaptation this is quite often neglected in favour of predicted sensory consequences based on efference copies of motor commands. Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences. We use a training paradigm combining robot controlled hand movements with rotated visual feedback that eliminates the possibility to update predicted sensory consequences (‘exposure training’), but still recalibrates proprioception, as well as a classic training paradigm with self-generated movements in another set of participants. After each kind of training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements (‘active localization’) as well as based on proprioception only with robot-generated movements (‘passive localization’). In the exposure training group, we find indistinguishable shifts in passive and active hand localization, but after classic training, active localization shifts more than passive, indicating a contribution from updated predicted sensory consequences. Both changes in open-loop reaches and hand localization are only slightly smaller after exposure training as compared to after classic training, confirming that proprioception plays a large role in estimating limb position and in planning movements, even after adaptation. (data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
---

@@

However, there is another explanation for shifts in hand localization after visuomotor adaptation. Our lab, and others, have shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014; @Cameron2012], and a comparable proprioceptive change is induced with force-field adaptation [@Ostry2016; @Ohashi2019]. As proprioception, like predicted sensory consequences, also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018]. Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012]. Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.

Here we intend to further examine the contribution of proprioception to hand localization, relative to the contribution of predicted sensory consequences. To do this, we use a passive training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013]. This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences. That is, exposure training to visuoproprioceptive discrepancy should not lead to updates of predicted sensory consequences. However, the discrepancy between vision and proprioception does drive proprioceptive recalibration. Thus, changes in any kind of hand localization after this type of training should be due to proprioceptive recalibration only. We use the same experimental protocol as before @Hart2016, where classic visuomotor training should have changed predicted sensory consequences along with proprioception. This allows comparing shifts in hand localization between the two different types of training, and a better assessment of the contributions of predicted sensory consequences and proprioception to hand localization.

@@

## Exposure training

@@

## No-cursor reaching

@@

@@

@@

>>>>

---
title: 'Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization
  After Passive Visuoproprioceptive Discrepancy Training'
author:
- affiliation:
  - YORK
  - MANSOURA
  name: Ahmed A. Mostafa
- affiliation: YORK
  corresponding: thartbm@gmail.com
  email: thartbm@gmail.com
  name: Bernard Marius 't Hart
- affiliation: YORK
  email: deniseh@yorku.ca
  name: Denise Y.P. Henriques
address:
- address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
    Canada
  code: YORK
- address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
  code: MANSOURA
output:
  word_document: default
  rticles::plos_article: default
  pdf_document: default
  html_document:
    df_print: paged
csl: plos.csl
fig_caption: yes
header-includes:
- \usepackage{threeparttable}
- \usepackage[table]{xcolor}
- \usepackage{graphicx}
- \usepackage{tabularx}
- \usepackage{longtable}
- \usepackage{siunitx}
bibliography: bibliography.bib
abstract: |
  An accurate estimate of limb position is necessary for movement planning, before and after motor learning. Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but in studies and theories on motor adaptation this is quite often neglected in favour of predicted sensory consequences based on efference copies of motor commands. Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences. We use a training paradigm combining robot controlled hand movements with rotated visual feedback that eliminates the possibility to update predicted sensory consequences (‘exposure training’), but still recalibrates proprioception, as well as a classic training paradigm with self-generated movements in another set of participants. After each kind of training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements (‘active localization’) as well as based on proprioception only with robot-generated movements (‘passive localization’). In the exposure training group, we find indistinguishable shifts in passive and active hand localization, but after classic training, active localization shifts more than passive, indicating a contribution from updated predicted sensory consequences. Both changes in open-loop reaches and hand localization are only slightly smaller after exposure training as compared to after classic training, confirming that proprioception plays a large role in estimating limb position and in planning movements, even after adaptation. (data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

@@

However, there is another explanation for shifts in hand localization after visuomotor adaptation. Our lab, and others, have shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014; @Cameron2012], and a comparable proprioceptive change is induced with force-field adaptation [@Ostry2016; @Ohashi2019]. Just like efference-based predictions, proprioception also informs us on the location of our limbs. To investigate recalibrated proprioception we have on occasion used a task that is very similar to hand localization [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018]. Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012]. Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.

Here we intend to further examine the contribution of proprioception to hand localization, relative to the contribution of predicted sensory consequences. To do this, we use a training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013]. This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences. That is, exposure training to visuoproprioceptive discrepancy should not lead to updates of predicted sensory consequences. However, the discrepancy between vision and proprioception does drive proprioceptive recalibration. Thus, changes in any kind of hand localization after this type of training should be due to proprioceptive recalibration only. We use the same experimental protocol as before @Hart2016, where classic visuomotor training should have changed predicted sensory consequences along with proprioception. This allows comparing shifts in hand localization between the two different types of training, and a better assessment of the contributions of predicted sensory consequences and proprioception to hand localization.

@@

### Exposure training

@@

### No-cursor reaching

@@

### Localization

@@

### Reach aftereffects

@@

### Generalization maxima
